Probabilistic Catalogs from Crowded Star Fields

### 2012-09-13 paper/ms.tex
Luminosity function parameters
Some writing

\title{Probabilistic Catalogs from Crowded Star Fields}
\author{Brendon J. Brewer$^{1, 2, ^*}$, David W. Hogg$^{3}$,
and Daniel Foreman-Mackey$^{3}$ \\
\\
\small
$^1$ Department of Physics, University of California, Santa Barbara,\\
\small
CA 93106, USA \\
\small
$^2$ Department of Statistics, The University of Auckland,\\
\small
Private Bag 92019, Auckland 1142, New Zealand \\
\small
$^3$ Center for Cosmology and Particle Physics, Department of Physics,
New York University,\\
\small
4 Washington Place, New York, NY, 10003, USA\\
\small
$^*$\texttt{bj.brewer@auckland.ac.nz}
}

@@

\subsection{The Sampling Distribution}
The sampling distribution is a probabilistic model for the process that
generates the data; it describes the probability distribution we would use
to predict the data if we happened to know the true catalog. Thus, we need
a prescription for simulating images from a catalog:
\begin{eqnarray}
\theta = \left\{N, \beta, \left\{x_i, y_i\right\}_{i=1}^N, 
\left\{f_i\right\}_{i=1}^N\right\}
\end{eqnarray}
If we knew the true catalog, we could compute the image we would expect to see,
in the absence of noise. This mock image is given by:
\begin{eqnarray}
\mathcal{M}(x, y) &=& \sum_{i=1}^N \mathcal{P}(x - x_i, y - y_i)
\end{eqnarray}
where $\mathcal{P}$ is the pixel-convolved point spread function (PSF).
Throughout this paper we
assume the pixel-convolved PSF is a known weighted mixture of two concentric
circular
Gaussians with widths $s_1$ and $s_2$:
\begin{eqnarray}
\mathcal{P}(x, y) = \frac{w}{2\pi s_1^2}\exp
\left[
-\frac{1}{2s_1^2}\left(x^2 + y^2\right)
\right]
+ \frac{1-w}{2\pi s_2^2}\exp
\left[-\frac{1}{2s_2^2}\left(x^2 + y^2\right)
\right].
\end{eqnarray}
The observed image is assumed to be generated from the mock image (evaluated
at the center of each pixel) plus Gaussian noise:
\begin{eqnarray}
y_{ij} &=& \mathcal{M}(x_{ij}, y_{ij}) + \epsilon_{ij}
\end{eqnarray}
where $\epsilon_{ij} \sim \mathcal{N}(0, \sigma^2)$.

\subsection{The Prior Distribution}
The prior distribution for the number of stars $N$ is uniform between 0 and
$N_{\rm max}$.

\section{Broken Power-Law Distribution}
The broken power-law distribution is defined by:
\begin{eqnarray}
p(x) &\propto&
\left\{
\begin{array}{lcr}
0, & & x < x_1 \\
x^{-\alpha_1 - 1} & & x_1 \leq x \leq x_2 \\
x^{-\alpha_2 - 1} & & x > x_2
\end{array}
\right.
.
\end{eqnarray}

@@

\section{MCMC Implementation}
\subsection{Current}
I am using Diffusive Nested Sampling \citep{dnest}, which is really just
Metropolis-Hastings but targeting a fancy distribution instead of the actual
posterior, and then assigning importance weights when you want the posterior.
Nested Sampling is good because it is invariant under monotonic transformations
of the likelihood function. i.e. the exploration only depends on things like
whether one point is better than another, and not on how much better it is.

@@

\appendix
\section{Pixel Convolved PSF}
Consider a ``true image'' $f_0(x, y)$ with infinite resolution. We now model how
this image gives rise to the observed data. First, it is convolved with a
PSF $g(\delta_x, \delta_y)$ (assumed normalized to 1) to give the blurred image:
\begin{eqnarray}
f_{\rm{blurred}}(x, y) &=& \iint f_0(x-\delta_x, y-\delta_y)
g(\delta_x, \delta_y) \, d\delta_x \, d\delta_y \label{blur}
\end{eqnarray}
Then, we observe the blurred image with a certain pixellation. Consider a pixel.
The true flux $F$ within the pixel is:
\begin{eqnarray}
F &=& \iint_{\rm pixel} f_{\rm blurred}(x, y) \,dx'\,dy'\\
&=& \iint f_{\rm blurred}(x, y)\mathbbm{1}\left[(x,y) \in \textnormal{pixel}
\right] \,dx\,dy \label{pixel}
\end{eqnarray}
Now consider $F$ as a function of the central location $(x_c, y_c)$ of the
pixel.

>>>>

\title{Probabilistic Catalogs from Crowded Star Fields}
\author{Brendon J. Brewer$^{1,*}$, David W. Hogg$^{2}$,
and Daniel Foreman-Mackey$^{2}$ \\
\\
\small
$^1$ Department of Statistics, The University of Auckland,\\
\small
Private Bag 92019, Auckland 1142, New Zealand \\
\small
$^2$ Center for Cosmology and Particle Physics, Department of Physics,
New York University,\\
\small
4 Washington Place, New York, NY, 10003, USA\\
\small
$^*$\texttt{bj.brewer@auckland.ac.nz}
}

@@

\subsection{The Sampling Distribution}
The sampling distribution is a probabilistic model for the process that
generates the data; it describes the probability distribution we would use
to predict the data if we happened to know the true catalog. In our case,
the data will be an $n \times n$ array of pixel intensities:
\begin{eqnarray}
\{Y_{ij}\}
\end{eqnarray}
The center of each pixel is:
\begin{eqnarray}
\{x_{ij}, y_{ij}\}.
\end{eqnarray}
Thus, we need
a prescription for simulating an image $\{Y_{ij}\}$ from a catalog
$\theta = \left\{N, \beta, \left\{x_i, y_i\right\}_{i=1}^N, 
\left\{f_i\right\}_{i=1}^N\right\}$.
If we knew the true catalog, we could compute the
``mock image'' we would expect to see
in the absence of noise. This mock image (at infinite resolution) is given by:
\begin{eqnarray}
\mathcal{M}(x, y) &=& \sum_{i=1}^N \mathcal{P}(x - x_i, y - y_i)
\end{eqnarray}
where $\mathcal{P}$ is the pixel-convolved point spread function (PSF).
Throughout this paper we
assume the pixel-convolved PSF is a known weighted mixture of two concentric
circular
Gaussians with widths $s_1$ and $s_2$:
\begin{eqnarray}
\mathcal{P}(x, y) = \frac{w}{2\pi s_1^2}\exp
\left[
-\frac{1}{2s_1^2}\left(x^2 + y^2\right)
\right]
+ \frac{1-w}{2\pi s_2^2}\exp
\left[-\frac{1}{2s_2^2}\left(x^2 + y^2\right)
\right].
\end{eqnarray}
The finite-resolution observed image is assumed to be generated from the mock
image (evaluated at the center of each pixel) plus Gaussian noise:
\begin{eqnarray}
Y_{ij} &=& \mathcal{M}(x_{ij}, y_{ij}) + \epsilon_{ij}
\end{eqnarray}
where the errors $\{\epsilon_{ij}\}$ are independent and normally distributed:
\begin{eqnarray}
\epsilon_{ij} \sim \mathcal{N}(0, \sigma^2).
\end{eqnarray}
Throughout this paper, we will assume that the noise variance $\sigma^2$ is
known. This is not usually true in practice (although it is often assumed).
However it is straightforward to extend the model so that $\sigma$ is an
unknown parameter.

\subsection{The Prior Distribution}
The prior distribution for the number of stars $N$ is uniform between 0 and
some maximum number $N_{\rm max}$. The extent of the image is assumed to be
from $x=-1$ to $x=1$ and from $y=-1$ to $y=1$ in arbitrary units, and the
positions of the stars are assigned independent uniform priors:
\begin{eqnarray}
x_i &\sim& \textnormal{Uniform}(-1.1, 1.1) \\
y_i &\sim& \textnormal{Uniform}(-1.1, 1.1) \\
\end{eqnarray}
The model for the luminosity function is a broken power-law with four
parameters:
\begin{eqnarray}
\beta = \{h_1, h_2, \alpha_1, \alpha_2\}.
\end{eqnarray}

\section{MCMC Implementation}
The sampling was implemented using the Diffusive Nested Sampling \citep{dnest}
method (hereafter DNS). DNS is based on the Metropolis-Hastings algorithm and
is very generally applicable. The main difference between DNS and the standard
Metropolis-Hastings algorithm is that the target distribution is modified.
Rather than simply exploring the posterior distribution over catalog space,
DNS constructs an alternative target distribution which is a mixture of the
prior distribution with more constrained versions of the prior distribution.
The modified target distribution assists

Nested Sampling is good because it is invariant under monotonic transformations
of the likelihood function. i.e. the exploration only depends on things like
whether one point is better than another, and not on how much better it is.

@@

@@

\appendix

\section{Broken Power-Law Distribution}
The broken power-law distribution is defined by:
\begin{eqnarray}
p(x) &\propto&
\left\{
\begin{array}{lcr}
0, & & x < x_1 \\
x^{-\alpha_1 - 1} & & x_1 \leq x \leq x_2 \\
x^{-\alpha_2 - 1} & & x > x_2
\end{array}
\right.
.
\end{eqnarray}
The free parameters of the broken power-law are:
\begin{eqnarray}
\beta = \{x_1, x_2, \alpha_1, \alpha_2\}
\end{eqnarray}

\section{Pixel Convolved PSF}
Consider a ``true image'' $f_0(x, y)$ with infinite resolution. We now model how
this image gives rise to the observed data. First, it is convolved with a
PSF $g(\delta_x, \delta_y)$ (assumed normalized to 1) to give the blurred image:
\begin{eqnarray}
f_{\rm{blurred}}(x, y) &=& \iint f_0(x-\delta_x, y-\delta_y)
g(\delta_x, \delta_y) \, d\delta_x \, d\delta_y \label{blur}
\end{eqnarray}
Then, we observe the blurred image with a certain pixellation. Consider a pixel.
The true flux $F$ within the pixel is:
\begin{eqnarray}
F &=& \iint_{\rm pixel} f_{\rm blurred}(x, y) \,dx'\,dy'\\
&=& \iint f_{\rm blurred}(x, y)\mathbbm{1}\left[(x,y) \in \textnormal{pixel}
\right] \,dx\,dy \label{pixel}
\end{eqnarray}
Now consider $F$ as a function of the central location $(x_c, y_c)$ of the
pixel.
